Week 1 Assignment 
=======================

## Loading and preprocessing the data
```{r}
library(ggplot2)
library(dplyr)
activity <- read.csv("activity.csv", colClasses = c("integer", "Date", "factor"))
activity$month <- as.numeric(format(activity$date, "%m"))
noNA <- na.omit(activity)
rownames(noNA) <- 1:nrow(noNA)
```


## What is the mean total number of steps taken per day?

Calculate the total number of steps taken per day:
```{r}
total.step <- aggregate(noNA$steps, list(Date = noNA$date), FUN = "sum")$x
```

Here is a histogram of the total number of steps taken each day:
```{r}
ggplot(noNA, aes(date, steps)) 
```

The mean and median of the total number of steps taken per day are:
```{r}
mean(total.step)
median(total.step)
```



## What is the average daily activity pattern?
Make a time series plot:
```{r}
avgSteps <- aggregate(noNA$steps, list(interval = as.numeric(as.character(noNA$interval))), FUN = "mean")
names(avgSteps)[2] <- "meanOfSteps"

ggplot(avgSteps, aes(interval, meanOfSteps)) + geom_line(color = "blue", size = 0.8) + labs(title = "Time Series of average number os steps taken")
```

Which 5-munite interval, on average across all the days in the dataset, contains the maximum number of steps:
```{r}
avgSteps[avgSteps$meanOfSteps == max(avgSteps$meanOfSteps), ]
```


## Imputing missing values
Calculate the total number of missing values in the dataset:
```{r}
sum(is.na(activity))
```

Create a new dataset with missing value filled with average of that interval:
```{r}
newData <- activity 
for (i in 1:nrow(newData)) {
    if (is.na(newData$steps[i])) {
        newData$steps[i] <- avgSteps[which(newData$interval[i] == avgSteps$interval), ]$meanOfSteps
    }
}
head(newData)
```

Make a histogram and report the mean and median total number of steps taken per day:
```{r}
ggplot(newData, aes(date, steps)) + geom_bar(stat = "identity",
                                             colour = "blue",
                                             fill = "blue",
                                             width = 0.6) + facet_grid(. ~ month, scales = "free") + labs(title = "Histogram of Total Number of Steps Taken Each Day (no missing data)", x = "Date", y = "Total number of steps")

newTotalSteps <- aggregate(newData$steps, list(Date = newData$date), FUN = "sum")$x
newMean <- mean(newTotalSteps)
newMean

newMedian <- median(newTotalSteps)
newMedian
```

Compare the newMean and newMedian with the two before:
```{r}
oldMean <- mean(total.step)
oldMedian <- median(total.step)

newMean - oldMean

newMedian - oldMedian
```
The new mean is the same as the old mean, while the new median is larger than the old median.


## Are there differences in activity patterns between weekdays and weekends?

Create a new factor indicating if a given day is a weekday or weekend:
```{r}
newData$weekdays <- factor(format(newData$date, "%A"))
levels(newData$weekdays)

levels(newData$weekdays) <- list(weekday = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday"), weekend = c("Saturday", "Sunday"))
levels(newData$weekdays)
table(newData$weekdays)
```

Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). 
```{r}

avgSteps <- aggregate(newData$steps, 
                      list(interval = as.numeric(as.character(newData$interval)), 
                           weekdays = newData$weekdays),
                      FUN = "mean")
names(avgSteps)[3] <- "meanOfSteps"
library(lattice)
xyplot(avgSteps$meanOfSteps ~ avgSteps$interval | avgSteps$weekdays, 
       layout = c(1, 2), type = "l", 
       xlab = "Interval", ylab = "Number of steps")
```